Monte Carlo tree search compared to A\(^*\) in airspace configuration decision problem

### 2018-05-03 article/article.tex
(+) Added state of the art on monte carlo
And low, high, normal and aircraft or \text-ed

@@

The simplest approch considers only the number of aircraft per sector $N_{aircraft}$. We
defined two arbitary thresholds, namely $th_{low}$ and $th_{high}$, defining
three zones where the workload is low (resp. normal and high) if $N_{aircraft} < th_{low}$
(resp. $th_{low} < N_{aircraft} < th_{high}$ and $th_{high} < N_{aircraft}$).
Those workload levels are translated to probabilities $p_{low}$, $p_{normal}$ and
$p_{high}$, where $p_{low}(S)$ is the probability that the sector $S$ is underloaded
(resp. balanced and overloaded). In short, for a sector $S$ at time $t$ and $N_{aircraft}$:
\begin{equation}
  \begin{aligned}
    p_{low}^{S, t} &= \mathbb{1}_{[0, th_{low}]}(N_{aircraft}) \\
    p_{normal}^{S, t} &= \mathbb{1}_{[th_{low}, th_{high}]}(N_{aircraft})\\
    p_{high}^{S, t} &= \mathbb{1}_{[th_{low}, \infty[}(N_{aircraft})
  \end{aligned}
\end{equation}
where $\mathbb{1}$ is the indicator function defined by:
\begin{equation}
  \mathbb{1}_{I}(x) = 
  \begin{cases}
    1 & \text{if\,} \quad x \in I\\
    0 & \text{otherwise}
  \end{cases}
\end{equation}

This expression of the workload, expressed in term of probabilities, allow us to
use a more complex model for the workload such as a neural network
(see~\cite{gianazza2010forecasting}). This method learns the probabilities $p_{low}$,
$p_{normal}$ and $p_{high}$ based on parameters including trafic complexity and
the complexity of the sector (e.g. airspace volume).

@@

In order to evaluate a given partition $P$ at time $t$, a cost $C(P, t)$ needs
to be defined. In this paper, this cost depends on the workload in each the
sector. The definitions given in~\cite{ferrari2017} are used to represent a high
(respectively normal and low) cost for partition \(P\):
\begin{equation}
  \begin{aligned}
    c_+(P, t) &= \sum_{S \in P_t} \delta_{h}(S, t) \cdot p_{high}^{S, t} \cdot |S|^2\\
    c_=(P, t) &= \left(
    \sum_{S \in P_t} \delta_{n}(S, t) \cdot p_{norm}^{S} \cdot |S|^{-2}
    \right)^{-1}\\
    c_-(P, t) &= \sum_{S \in P_t} \delta_{l}(S, t) \cdot p_{low}^{S, t} \cdot |S|^{-2}
  \end{aligned}
\end{equation}
with $\delta_{h}(P, t)$ (resp. $\delta_{n}(P_t)$ and $\delta_{l}(P_t)$) equals 1
if the probability $p_{high}$ (resp. $p_{normal}$ and $p_{low}$) is superior to
the two others, and 0 otherwise.

>>>>

The Monte Carlo tree search algorithm has been quickly used in two players game.
Only three years after its apparition in 1990, Br\"ugmann applies it to the Go
game in~\cite{brugmann1993mcgo}. While the Monte-Carlo tree search is still
extensively used in two players games (and especially the Go game), its
adaptation to one player game has been worked on. Auer \textit{et al.} propose
an
upper confidence bound formula which appears to be more efficient for one player
games as seen in~\cite{sebag2010fuse}. The introduced formula uses the standard
deviation of the outcome of the simulations.
The latter article also uses the rapid
action value estimation (RAVE) technique to quicken the convergence of the
algorithm. RAVE uses the ``all moves as first'' heuristic, in which all
moves\footnote{a move is informally considered as a decision taken regarding
which state to choose while descending the search tree}
seen during simulations are considered as a first move. This allows the
algorithm to update more statistics in one simulation. Gelly \textit{et al.} use
in~\cite{gelly2012grand} the RAVE technique coupled with several heuristics. The
heuristics bias the initialisaton of a node in the search tree, pre filling its
statistics using prior knowledge of the problem.

@@

The simplest approch considers only the number of aircraft per sector $N_\text{aircraft}$. We
defined two arbitary thresholds, namely $th_\text{low}$ and $th_\text{high}$, defining
three zones where the workload is low (resp. normal and high) if $N_\text{aircraft} < th_\text{low}$
(resp. $th_\text{low} < N_\text{aircraft} < th_\text{high}$ and $th_\text{high} < N_\text{aircraft}$).
Those workload levels are translated to probabilities $p_\text{low}$, $p_\text{normal}$ and
$p_\text{high}$, where $p_\text{low}(S)$ is the probability that the sector $S$ is underloaded
(resp. balanced and overloaded). In short, for a sector $S$ at time $t$ and $N_\text{aircraft}$:
\begin{equation}
  \begin{aligned}
    p_\text{low}^{S, t} &= \mathbb{1}_{[0, th_\text{low}]}(N_\text{aircraft}) \\
    p_\text{normal}^{S, t} &= \mathbb{1}_{[th_\text{low}, th_\text{high}]}(N_\text{aircraft})\\
    p_\text{high}^{S, t} &= \mathbb{1}_{[th_\text{low}, \infty[}(N_\text{aircraft})
  \end{aligned}
\end{equation}
where $\mathbb{1}$ is the indicator function defined by:
\begin{equation}
  \mathbb{1}_{I}(x) =
  \begin{cases}
    1 & \text{if\,} \quad x \in I\\
    0 & \text{otherwise}
  \end{cases}
\end{equation}

This expression of the workload, expressed in term of probabilities, allow us to
use a more complex model for the workload such as a neural network
(see~\cite{gianazza2010forecasting}). This method learns the probabilities $p_\text{low}$,
$p_\text{normal}$ and $p_\text{high}$ based on parameters including trafic complexity and
the complexity of the sector (e.g. airspace volume).

@@

In order to evaluate a given partition $P$ at time $t$, a cost $C(P, t)$ needs
to be defined. In this paper, this cost depends on the workload in each the
sector. The definitions given in~\cite{ferrari2017} are used to represent a high
(respectively normal and low) cost for partition \(P\):
\begin{equation}
  \begin{aligned}
    c_+(P, t) &= \sum_{S \in P_t} \delta_{h}(S, t) \cdot p_\text{high}^{S, t} \cdot |S|^2\\
    c_=(P, t) &= \left(
    \sum_{S \in P_t} \delta_{n}(S, t) \cdot p_{norm}^{S} \cdot |S|^{-2}
    \right)^{-1}\\
    c_-(P, t) &= \sum_{S \in P_t} \delta_{l}(S, t) \cdot p_\text{low}^{S, t} \cdot |S|^{-2}
  \end{aligned}
\end{equation}
with $\delta_{h}(P, t)$ (resp. $\delta_{n}(P_t)$ and $\delta_{l}(P_t)$) equals 1
if the probability $p_\text{high}$ (resp. $p_\text{normal}$ and $p_\text{low}$) is superior to
the two others, and 0 otherwise.
